---
title: "Machine Learning Project"
author: "Miguel Alomar"
date: "Wednesday, June 17th 2015"
output: html_document
---

##Overview

This presentation for the Coursera class "Practice Machine Learning" will show how I applied
a prediction model to predict the manner in which a fitness exercise is done.
It will go through the following steps:
* Loading
* Selecting the variables to use
* Visualizations
* Creating model
* Predicting

##Loading the data

Loading the data is done as usual.
```{r}
library(caret);
library(kernlab);
library("ggplot2", lib.loc="C:/Program Files/R/R-3.1.1/library")
pml <- read.csv("C:/Personal/coursera/Machine_learning/pml-training.csv", sep=",",
                na.strings = c("NA","#DIV/0!"))
pmltest <- read.csv("C:/Personal/coursera/Machine_learning/pml-testing.csv", sep=",",
                    na.strings = c("NA","#DIV/0!"))
```




---
##Selecting variables

The datasets have 160 variables, not all of them are useful or have values.
These are the variables I selected, to decide which ones to use I discarded the ones that looked
like identifiers or time variables, then in a second step I selected the numeric variables that 
didn't have NA values.

```{r}
set.seed(32343)
SmallPml <- pml[,c("roll_belt","pitch_belt","yaw_belt","total_accel_belt",
                   "gyros_belt_x","gyros_belt_y","gyros_belt_z",
                   "accel_belt_x","accel_belt_y","accel_belt_z",
                   "magnet_belt_x","magnet_belt_y","magnet_belt_z",
                   "roll_arm","pitch_arm","yaw_arm",
                   "total_accel_arm",
                   "gyros_arm_x","gyros_arm_y","gyros_arm_z",
                   "accel_arm_x","accel_arm_y","accel_arm_z",
                   "magnet_arm_x","magnet_arm_y","magnet_arm_z",
                   "roll_dumbbell","pitch_dumbbell","yaw_dumbbell",
                   "classe"       )              ]
SmallTest <- pmltest[,c("roll_belt","pitch_belt","yaw_belt","total_accel_belt",
                   "gyros_belt_x","gyros_belt_y","gyros_belt_z",
                   "accel_belt_x","accel_belt_y","accel_belt_z",
                   "magnet_belt_x","magnet_belt_y","magnet_belt_z",
                   "roll_arm","pitch_arm","yaw_arm",
                   "total_accel_arm",
                   "gyros_arm_x","gyros_arm_y","gyros_arm_z",
                   "accel_arm_x","accel_arm_y","accel_arm_z",
                   "magnet_arm_x","magnet_arm_y","magnet_arm_z",
                   "roll_dumbbell","pitch_dumbbell","yaw_dumbbell" )]


```

---
##Visualizations
I created several plots to try to find patters or useful variables but it didn't help much.

```{r}
featurePlot(x=SmallPml[c("gyros_belt_x","gyros_belt_y","gyros_belt_z")],
            y=SmallPml$classe,
            plot="pairs"
            )
```

##Creating model
To create the model I selected only 25% of the data in my training set, I did that to get
around the memory limitations of my computer, plus I found that 25% was enough to get good
predictions.
I did try with several methods. rf: Random forest, rpart: Predict tree, gbm:adaboost and few more.
The one that gave me best results was Random forest so that is the one I went with.

```{r}
inTrain <- createDataPartition(y=SmallPml$classe,p=0.25, list=FALSE);
training <- SmallPml[inTrain,]
testing  <- SmallPml[-inTrain,]
dim(training)
modFit <- train(classe ~ .,
                data=training,
                method="rf",
                prox=TRUE)
modFit
plot(modFit)
```



---
##Predicting
Now we use predict function to see how accurate our model is.
The Accuracy was around 93.5%

```{r}
testing <- testing[complete.cases(testing),]
pred <- predict(modFit,testing)
testing$predRight <- pred==testing$classe
table(pred,testing$classe)
```

and finally we using against the pmltest data to generate the 20 predictions.
Sure enough, I got 19 out of 20 cases correct, that is 95% of the cases.
```{r}
pred <- predict(modFit,SmallTest)
```
